---
title: "EAR Analysis"
author: "J. Lucas McKay"
date: "December 1 2021"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)
```

# Exploratory data analysis

```{r}
library(tidyverse)
library(motionTools)

# "plus equals" function
`%+=%` = function(e1,e2) eval.parent(substitute(e1 <- e1 + e2))

  # rename(emg_baseline = `EMG Baseline`, emg_peak = `EMG Peak`, emg_peak_base = `Peak - Base`) |> 

data = read_csv("EAR_PD_Table.csv") |> 
  rename(`EMG Peak - Base` = `Peak - Base`) |> 
  mutate(Muscle = case_when(
    Muscle %contains% "MGAS" ~ "MGAS",
    Muscle %contains% "TA" ~ "TA",
    Muscle %contains% "SOL" ~ "SOL",
    TRUE ~ ""
  )) |> 
  filter(Muscle != "")

# Consider only the shortening reactions
data2 = data |> 
  filter((Direction == 90 & Muscle == "MGAS")|(Direction == 90 & Muscle == "SOL")|(Direction == 270 & Muscle == "TA"))
```

# EMG levels during shortening vs. condition

```{r}
data3 = data2 |> 
  select(-Direction) |> 
  pivot_longer(starts_with("EMG"), names_to = "Window", values_to = "EMG") |> 
  mutate(Condition = factor(Condition))
```

```{r}
data3 |>
  filter(Window == "EMG Baseline") |> 
  ggplot(aes(EMG,fill = Condition)) + geom_density(alpha = 0.5) + facet_grid(cols = vars(State),rows=vars(Muscle)) +
  scale_fill_viridis_d(end = 0.6) +
  theme_minimal() +
  labs(title = "EMG Baseline")
```

```{r}
data3 |>
  filter(Window == "EMG Peak - Base") |> 
  ggplot(aes(EMG,fill = Condition)) + geom_density(alpha = 0.5) + facet_grid(cols = vars(State),rows=vars(Muscle)) +
  scale_fill_viridis_d(end = 0.6) +
  theme_minimal() +
  labs(title = "EMG Peak - Base")
```


# Average values

```{r}
emg_summary = data3 |> group_by(State,Condition,Muscle,Window) |> summarize(EMG = mean(EMG, na.rm = T), .groups = "drop")
```

```{r}
emg_summary |> 
  filter(Window == "EMG Baseline") |> 
  ggplot(aes(Condition,EMG,fill = Condition)) +
    geom_col() +
    facet_grid(cols = vars(State),rows=vars(Muscle)) +
  scale_fill_viridis_d(end = 0.6) +
  theme_minimal() +
  labs(title = "EMG Baseline")
```

```{r}
emg_summary |> 
  filter(Window == "EMG Peak - Base") |> 
  ggplot(aes(Condition,EMG,fill = Condition)) +
    geom_col() +
    facet_grid(cols = vars(State),rows=vars(Muscle)) +
  scale_fill_viridis_d(end = 0.6) +
  theme_minimal() +
  labs(title = "EMG Peak - Base")
```

# Effects of testing condition on baseline EMG

```{r}
emg_baseline = data3 |>
  filter(Window == "EMG Baseline")

mdl_baseline = lmerTest::lmer(EMG~Condition+State+Muscle+(1|Name), data = emg_baseline)
summary(mdl_baseline)
anova(mdl_baseline)
```

# Effects of testing condition on peak EMG

```{r}
emg_peak_baseline = data3 |>
  filter(Window == "EMG Peak - Base")
mdl_peak_baseline = lmerTest::lmer(EMG~Condition+State+Muscle+(1|Name), data = emg_peak_baseline)
summary(mdl_peak_baseline)
anova(mdl_peak_baseline)
```

# Methods

Average EMG levels during baseline ("Baseline," X-X ms prior to perturbation onset) and during the shortening reaction ("Peak," X-X ms) were calculated for each muscle in each patient. Baseline values were subtracted from Peak values prior to analysis.
Variation in EMG level with medication state and testing condition were assessed with separate linear mixed models for Baseline and Peak.

# Results

Linear mixed models (Condition x State x Muscle) identified significant effects of Condition (P<0.001) on baseline EMG but not on peak EMG (P=0.37), as well as highly significant effects of medication state in both baseline (P<<0.001) and peak (P<<0.001).


